Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-23 article/article.Rmd
Commit 2024-06-23 17:12

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal variability of the number of CE events normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have been also normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values and the solid circles are outliers."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-MW-1.png")
```
<!--
```{r echo=FALSE, fig.cap="Distribution of the number of CE cases"}
knitr::include_graphics("../images/climCEmonth-1.png")
```
--->

<!-- ```{r echo=FALSE, fig.cap="Distribution of CE above reference"} -->
<!-- knitr::include_graphics("../images/relative_distributions-2.png") -->
<!-- ``` -->
<!--
```{r echo=FALSE, fig.cap="Distribution of CE above reference"}
knitr::include_graphics("../images/relative_distributions-4.png")
```
-->

The distribution of the number of CE events as a function of OIR is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently, there is an inverse
relationship between the frequency of CE events and OIR with an exponential-like
decline. This is expected, as the stronger the CE events are, the rarer the
conditions favoring the occurrence of CE events. For the majority (over $62\,\%$) of
the CE events the OIR is below the long-term average of
$`r round(mean(ST_E_yearly$GLB_diff.mean), 1)`\,\text{W}/\text{m}^2$,
while about $8.1\,\%$ of the events correspond to OIR larger than
$100\,\text{W}/\text{m}^2$ and up to the highest value of $412.4\,\text{W}/\text{m}^2$.
This distribution is indicative of the magnitude and the probability of the expected
CE events over Thessaloniki.  Similar distribution of CE events, albeit with larger
OIR values, has been reported by @Vamvakas2020, for the city of Patras. This site is
located $2.5^\circ$ south of Thessaloniki and is exposed to air masses coming mainly
from the eastern Mediterranean resulting in different cloud patterns, that may affect
the characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several minutes or even
more than an hour. In order to study the characteristics of these consecutive events,
we grouped them into bins of increasing duration in steps of one minute.

We have identified `r ST_G0[, .N]` groups of CE in the
whole period of study, where the
group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.

Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}) shows the frequency
distribution of the CE events according to their duration.  We conclude that although
some groups of events last for more than an hour about $80\,\%$ have duration of less
than 5 minutes.

@@

The relation between the duration of the CE groups and the excess irradiation of the
groups has also been investigated (Figure\nobreakspace{}\ref{fig:group-2d}).
Evidently, evens of high excess irradiation have small duration and vice versa. The
vast majority of grouped events are associated with small excess irradiation (e.g.
$<5\,\text{kJ}/\text{m}^2$) and small duration (e.g. $<5\,\text{min}$) while groups
with strong excess irradiation and long duration are rare.  Similar results on this
relation, have been reported by @Zhang2018, on a study using a far higher sampling
rate than ours.

@@

<!-- ```{r echo=FALSE, fig.cap="Relation of mean over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-4.png") -->
<!-- ``` -->
<!-- ```{r echo=FALSE, fig.cap="Relation of maximun over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-8.png") -->
<!-- ``` -->

@@

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Seasonal statistics of the number of ECE occurancies for each month, normalized to the maximum occurancies on March. The box represents the values of the low $25\\,\\%$ percentile to $75\\,\\%$ percentile, where the thick horizontal line inside is the mean, the vertical lines extend to the max imum and minimum vales, the dots are outlier values, and the rhombus is the mean."}
knitr::include_graphics("../images/clim-ECE-month-norm-MAX-median-N-2.png")
```

@@

<!-- -   SZA -->

<!-- ### Yearly plots relative to all GHI data. -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-1.png") -->
<!-- ``` -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-2.png") -->
<!-- ``` -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-3.png") -->
<!-- ``` -->

@@

# Discussion and conclusions

@@

The magnitude of the ECE events doesn't exceeds the values reported from sites with
more favourable conditions for the phenomenon. Although, the climatological
characteristic of the ECE events, showed that the most active months are spread on
half of the year (March -- June and September and October).
We found that CE conditions, can have a duration of more than an hour in rare
cases, with the bulk of the cases having a duration under 5 minutes.
Some of the characteristics of CE and ECE events we analysed,
have strong similarities with results
by @Vamvakas2020, for a city southern of Thessaloniki, with the analogues differences on the
 intensity of the solar radiation.

>>>>

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal variability of the number of CE events normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have been also normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values and the solid circles are outliers."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-MW-1.png")
```

<!--
```{r echo=FALSE, fig.cap="Distribution of the number of CE cases"}
knitr::include_graphics("../images/climCEmonth-1.png")
```
--->

<!-- ```{r echo=FALSE, fig.cap="Distribution of CE above reference"} -->
<!-- knitr::include_graphics("../images/relative_distributions-2.png") -->
<!-- ``` -->

<!--
```{r echo=FALSE, fig.cap="Distribution of CE above reference"}
knitr::include_graphics("../images/relative_distributions-4.png")
```
-->

The distribution of the number of CE events as a function of OIR is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently, there is an inverse
relationship between the frequency of CE events and OIR with an exponential-like
decline. This is expected, as the stronger the CE events are, the rarer the
conditions favoring the occurrence of CE events. For the majority (over $62\,\%$) of
the CE events the OIR is below the long-term average of
$`r round(mean(ST_E_yearly$GLB_diff.mean), 1)` \, \text{W} / \text{m}^2$, 
while about $8.1\,\%$ of the events correspond to OIR larger than
$100\,\text{W}/\text{m}^2$ and up to the highest value of $412.4\,\text{W}/\text{m}^2$.
This distribution is indicative of the magnitude and the probability of the expected
CE events over Thessaloniki.  Similar distribution of CE events, albeit with larger
OIR values, has been reported by @Vamvakas2020, for the city of Patras. This site is
located $2.5^\circ$ south of Thessaloniki and is exposed to air masses coming mainly
from the eastern Mediterranean resulting in different cloud patterns, that may affect
the characteristics and magnitude of the CE events.

@@

The duration of the CE events is variable and can last for several minutes or even
more than an hour. In order to study the characteristics of these consecutive events,
we grouped them into bins of increasing duration in steps of one minute.  We have
identified `r ST_G0[, .N]` groups of CE in the whole period of study, where the group
of the longest duration of
`r ST_G0[, max(GLB_diff.N)]` minutes occurred on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
Figure\nobreakspace{}\ref{fig:ceg-duration-distribution} shows the frequency
distribution of the CE events according to their duration.  We conclude that although
some groups of events last for more than an hour about $80\,\%$ have duration of less
than 5 minutes.

@@

The relation between the duration and the mean OIR of the groups has also been
studied (Figure\nobreakspace{}\ref{fig:group-2d}).  Evidently, evens of high excess
irradiation have small duration and vice versa. The vast majority of grouped events
are associated with small excess irradiation (e.g. $<5\,\text{kJ}/\text{m}^2$) and
small duration (e.g. $<5\,\text{min}$) while groups with strong excess irradiation
and long duration are very rare.  Similar results of this relation have been
reported by @Zhang2018, on a study using a far higher sampling rate than ours.

@@

<!-- ```{r echo=FALSE, fig.cap="Relation of mean over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-4.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Relation of maximun over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-8.png") -->
<!-- ``` -->

@@

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Seasonal variability of the number of ECE events normalized to the maximum occurring in March, in the form of a box and whisker plot. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles. The thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-ECE-month-norm-MAX-median-N-2.png")
```

@@

<!--  SZA -->

<!-- ### Yearly plots relative to all GHI data. -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-1.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-2.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-3.png") -->
<!-- ``` -->

@@

# Conclusions

@@

The magnitude of the ECE events does not exceeds the values reported for sites with
more favourable conditions for the phenomenon. The climatological characteristic of
the ECE events, showed that the most active months are spread on half of the year
(March -- June and September and October).  We found that CE conditions, can have a
duration of more than an hour in rare cases, with the bulk of the cases having a
duration under 5 minutes.  Some of the characteristics of CE and ECE events we
analysed,
have strong similarities with results
by @Vamvakas2020, for a city southern of Thessaloniki, with the analogues differences on the
 intensity of the solar radiation.
